---
title: "Digital request themes (LLM metadata, version 3)"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true
    number-sections: true
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(ggplot2)
library(scales)
library(flextable)

res <- readRDS(
  here::here(
    "02_data_output",
    "dig_req_llm_metadata_2.rds"
  )
)

flat <- res |>
  mutate(
    summary = map_chr(
      metadata_v3,
      ~ .x$summary %||% NA_character_
    ),
    business_area = map_chr(
      metadata_v3,
      ~ .x$business_area %||% NA_character_
    ),
    themes = map(
      metadata_v3,
      ~ unlist(.x$themes)
    )
  )

themes <- flat |>
  select(
    request_row_id,
    digital_request_number,
    business_area,
    themes
  ) |>
  unnest(themes) |>
  distinct(digital_request_number, themes, .keep_all = TRUE)

# Standardise capitalisation
themes_clean <- themes |>
  mutate(
    themes = str_squish(themes),
    themes = str_to_title(themes)
  )

# create flat metadata table for later use in report
flat_metadata <- res |>
    select(
        request_row_id,
        digital_request_number,
        metadata_v3
    ) |>
    unnest_wider(metadata_v3)

# Flextable ---------------

theme_ft <- function(x) {
  x |>
  autofit() |>
  theme_booktabs()
  }

# DT:datatable ------------------------------------------------------------

# Use DT:datatable to embed an interactive table with export options in an R session or markdown document

# copied from github: jbisanz/qiime2R/R/interactive_table.R
# https://rstudio.github.io/DT/options.html ... "DataTables has a large number of initialization options, which make it very flexible to customize the table."

interactive_table <- function(table, 
                              nrow = 5, 
                              not_as_factor = c("mrn", 
                                                "patient"
                                                ),
                              autoWidth = TRUE
) {
  # mutate(across(where(is.character) using forcats::as_factor, avoiding items in "not_as_factor"
  table <- table |> 
    mutate(across(where(is.character) 
                  & !matches(not_as_factor), 
                  forcats::as_factor))
  
  # configure table settings
  dtable <- DT::datatable(table, 
                          extensions = 'Buttons', 
                          filter = "top",
                          options = list(
                            pageLength = nrow, #nrow dictates pageLength,
                            #lengthMenu = c(5, 10, 15, 20),
                            # initComplete = JS(
                            #   "function(settings, json) {",
                            #   "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});", #we use the initComplete callback function in options to change the background color of the table header to black after the initialization:
                            #   "}"
                            dom = 'Bfrtip', 
                            buttons = c('copy',
                                        'csv',
                                        #'pdf',
                                        'excel'
                                        ),
                            autoWidth = autoWidth,
                            searchHighlight = TRUE)
  )
  
  return(dtable)
}


```

# Overview

This report summarises themes identified from digital requests using locally-hosted LLM-generated metadata.

The report is intended to support prioritisation and portfolio review by:

* identifying common request themes
* highlighting potentially related or duplicate requests
* summarising requests by business area

## Summary statistics

```{r}
summary_tbl <- tibble(
  Metric = c(
    "Requests analysed",
    "Business areas",
    "Themes"
  ),
  Value = c(
    nrow(flat),
    n_distinct(flat$business_area, na.rm = TRUE),
    n_distinct(themes$themes, na.rm = TRUE)
  )
)

summary_tbl |>
  flextable() |>
  theme_ft()
```

# Most common themes

## Word cloud

```{r}
#| out-width: 100%
#| fig-width: 16
#| fig-height: 12

theme_counts <- themes |>
  count(themes, sort = TRUE) |>
  #slice_head(n = 75) |> 
  filter(n >= 2)

theme_counts |>
  slice_head(n = 75) |>
  ggplot(
    aes(
      label = themes,
      size = n,
      colour = n
    )
  ) +
  ggwordcloud::geom_text_wordcloud_area(
    eccentricity = 0.5
  ) +
  scale_size_area(
    max_size = 60
  ) +
  scale_colour_viridis_c() +
  theme_void() +
  theme(
    legend.position = "none"
  )
```

## Top 20 themes

```{r}
theme_counts |>
  slice_head(n = 20) |>
  ggplot(
    aes(
      x = reorder(themes, n),
      y = n
    )
  ) +
  geom_col() +
  coord_flip()
```


# Potentially related requests

Requests that share multiple themes may represent:

* similar problems
* opportunities for collaboration
* duplicate requests
* opportunities to consolidate work

```{r}
pairs <- inner_join(
  themes,
  themes,
  by = "themes",
  relationship = "many-to-many"
) |>
  filter(
    digital_request_number.x < digital_request_number.y
  ) |>
  group_by(
    digital_request_number.x,
    digital_request_number.y
  ) |>
  summarise(
    shared_themes = paste(
      sort(unique(themes)),
      collapse = "; "
    ),
    n_shared_themes = n(),
    .groups = "drop"
  ) |>
  filter(n_shared_themes >= 2) |>
  arrange(
    desc(n_shared_themes),
    digital_request_number.x,
    digital_request_number.y
  )

pairs |>
    interactive_table(nrow = 5, 
                                      not_as_factor = c("mrn", 
                                                        "patient"
                                                        ),
                                ) 
```

# Request summaries

```{r}
flat |>
  transmute(
    digital_request_number,
    business_area,
    themes = map_chr(
      themes,
      paste,
      collapse = "; "
    ),
    summary = str_trunc(
      summary,
      width = 1000
    )
  ) |>
  arrange(
    business_area,
    digital_request_number
  ) |>
  interactive_table(nrow = 5, 
                                      not_as_factor = c("themes", 
                                                        "summary"
                                                        ),
                                ) 
```

# Appendix

## Heatmap: Themes x Business area

```{r}
theme_business <- themes |>
  count(
    business_area,
    themes
  )

ggplot(
  theme_business,
  aes(
    x = themes,
    y = business_area,
    fill = n
  )
) +
  geom_tile() +
  scale_fill_viridis_c() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )
```

<!-- ## Business area frequencies -->

<!-- ```{r} -->
<!-- flat |> -->
<!--   count( -->
<!--     business_area, -->
<!--     sort = TRUE -->
<!--   ) |> -->
<!--   rename( -->
<!--     Requests = n -->
<!--   ) |> -->
<!--   flextable() |> -->
<!--   theme_ft() -->
<!-- ``` -->





