---
title: "Digital request themes (LLM metadata, version 3)"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true
    number-sections: true
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(ggplot2)
library(scales)
library(flextable)

res <- readRDS(
  here::here(
    "02_data_output",
    "dig_req_llm_metadata_2.rds"
  )
)

flat <- res |>
  mutate(
    summary = map_chr(
      metadata_v3,
      ~ .x$summary %||% NA_character_
    ),
    business_area = map_chr(
      metadata_v3,
      ~ .x$business_area %||% NA_character_
    ),
    themes = map(
      metadata_v3,
      ~ unlist(.x$themes)
    ),
    n_findings_v2 = lengths(findings_v2),
    n_findings_v3 = lengths(findings_v3)
  )

themes <- flat |>
  select(
    request_row_id,
    digital_request_number,
    business_area,
    themes
  ) |>
  unnest(themes)

theme_ft <- function(x) {
  x |>
    theme_booktabs() |>
    autofit()
}
```

# Overview

This report summarises themes identified from digital requests using locally-hosted LLM-generated metadata.

The report is intended to support prioritisation and portfolio review by:

* identifying common request themes
* highlighting potentially related or duplicate requests
* summarising requests by business area
* identifying records with quality assurance findings

## Summary statistics

```{r}
summary_tbl <- tibble(
  Metric = c(
    "Requests analysed",
    "Business areas",
    "Themes",
    "Requests with QA findings"
  ),
  Value = c(
    nrow(flat),
    n_distinct(flat$business_area, na.rm = TRUE),
    n_distinct(themes$themes, na.rm = TRUE),
    sum(flat$n_findings_v2 > 0 | flat$n_findings_v3 > 0)
  )
)

summary_tbl |>
  flextable() |>
  theme_ft()
```

# Most common themes

## Theme frequencies

```{r}
theme_counts <- themes |>
  count(themes, sort = TRUE) |>
  mutate(
    percentage = percent(n / sum(n), accuracy = 0.1)
  )

theme_counts |>
  flextable() |>
  theme_ft()
```

## Top 15 themes

```{r}
theme_counts |>
  slice_head(n = 15) |>
  ggplot(
    aes(
      x = reorder(themes, n),
      y = n
    )
  ) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Most common themes",
    x = NULL,
    y = "Number of requests"
  )
```

# Themes by business area

The table below shows the most frequent themes within each business area.

```{r}
themes |>
  count(
    business_area,
    themes,
    sort = TRUE
  ) |>
  group_by(business_area) |>
  slice_head(n = 5) |>
  ungroup() |>
  rename(
    `Business area` = business_area,
    Theme = themes,
    Requests = n
  ) |>
flextable() |>
theme_ft()
```

# Potentially related requests

Requests that share multiple themes may represent:

* similar problems
* opportunities for collaboration
* duplicate requests
* opportunities to consolidate work

```{r}
pairs <- inner_join(
  themes,
  themes,
  by = "themes",
  relationship = "many-to-many"
) |>
  filter(
    digital_request_number.x < digital_request_number.y
  ) |>
  group_by(
    digital_request_number.x,
    digital_request_number.y
  ) |>
  summarise(
    shared_themes = paste(
      sort(unique(themes)),
      collapse = "; "
    ),
    n_shared_themes = n(),
    .groups = "drop"
  ) |>
  filter(n_shared_themes >= 2) |>
  arrange(
    desc(n_shared_themes),
    digital_request_number.x,
    digital_request_number.y
  )

pairs |>
  flextable() |>
  width(j = "shared_themes", width = 4) |> #optional
  theme_ft()
```

# Request summaries

```{r}
flat |>
  transmute(
    digital_request_number,
    business_area,
    themes = map_chr(
      themes,
      paste,
      collapse = "; "
    ),
    summary = str_trunc(
      summary,
      width = 1000
    )
  ) |>
  arrange(
    business_area,
    digital_request_number
  ) |>
  flextable() |>
  width(j = "summary", width = 6) |> #optional
  theme_ft()
```

# Quality assurance findings

The records below have one or more quality assurance findings generated during metadata extraction.

```{r}
qa <- flat |>
  transmute(
    digital_request_number,
    n_findings = n_findings_v2 + n_findings_v3,
    findings = pmap_chr(
      list(findings_v2, findings_v3),
      \(v2, v3) {
        c(v2, v3) |>
          unique() |>
          paste(collapse = "\n\n• ")
      }
    )
  ) |>
  mutate(
    findings = paste0("• ", findings)
  ) |>
  filter(n_findings > 0) |>
  arrange(desc(n_findings))

qa |>
  flextable() |>
  width(j = "findings", width = 8) |>
  valign(valign = "top") |>
  align(j = "n_findings", align = "center") |>
  theme_ft()
```

# Appendix

## Business area frequencies

```{r}
flat |>
  count(
    business_area,
    sort = TRUE
  ) |>
  rename(
    Requests = n
  ) |>
  flextable() |>
  theme_ft()
```

## All themes

```{r}
theme_counts |>
  arrange(desc(n)) |>
  flextable() |>
  theme_ft()
```
