---
title: "Digital request themes (LLM metadata, version 3)"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true
    number-sections: true
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(ggplot2)
library(scales)
library(flextable)

res <- readRDS(
  here::here(
    "02_data_output",
    "dig_req_llm_metadata_2.rds"
  )
)

flat <- res |>
  mutate(
    summary = map_chr(
      metadata_v3,
      ~ .x$summary %||% NA_character_
    ),
    business_area = map_chr(
      metadata_v3,
      ~ .x$business_area %||% NA_character_
    ),
    themes = map(
      metadata_v3,
      ~ unlist(.x$themes)
    ),
    n_findings_v2 = lengths(findings_v2),
    n_findings_v3 = lengths(findings_v3)
  )

themes <- flat |>
  select(
    request_row_id,
    digital_request_number,
    business_area,
    themes
  ) |>
  unnest(themes) |>
  distinct(digital_request_number, themes, .keep_all = TRUE)

# Standardise capitalisation
themes_clean <- themes |>
  mutate(
    themes = str_squish(themes),
    themes = str_to_title(themes)
  )

# Create groupings
theme_group = case_when(

  str_detect(themes, regex("AI|Artificial Intelligence")) ~
    "AI",

  str_detect(themes, regex(
    "Radiology|Imaging|PACS|RIS|Ultrasound|MRI",
    TRUE
  )) ~
    "Radiology & Imaging",

  str_detect(themes, regex(
    "Laboratory|LIMS|LIS|GLIMS|Pathology|Genetics",
    TRUE
  )) ~
    "Laboratory Systems",

  str_detect(themes, regex(
    "Pharmacy|Medication|Prescribing|Drug",
    TRUE
  )) ~
    "Pharmacy",

  str_detect(themes, regex(
    "Data|Database|Reporting|Dashboard|Governance",
    TRUE
  )) ~
    "Data & Information",

  str_detect(themes, regex(
    "Integration|Interoperability|Interface|SSO",
    TRUE
  )) ~
    "Integration",

  str_detect(themes, regex(
    "Migration|Replacement|Upgrade|Modernisation|Modernization",
    TRUE
  )) ~
    "System Change",

  str_detect(themes, regex(
    "Workflow|Process|Automation|Efficiency",
    TRUE
  )) ~
    "Workflow",

  str_detect(themes, regex(
    "Infrastructure|Hosting|Cloud|VM|Network|Hardware",
    TRUE
  )) ~
    "Infrastructure",

  str_detect(themes, regex(
    "Patient|Outpatient|Referral|Care Pathway",
    TRUE
  )) ~
    "Patient Pathways",

  TRUE ~ "Other"
)

# Flextable ---------------

theme_ft <- function(x) {
  x |>
  autofit() |>
  theme_booktabs()
  }

# DT:datatable ------------------------------------------------------------

# Use DT:datatable to embed an interactive table with export options in an R session or markdown document

# copied from github: jbisanz/qiime2R/R/interactive_table.R
# https://rstudio.github.io/DT/options.html ... "DataTables has a large number of initialization options, which make it very flexible to customize the table."

interactive_table <- function(table, 
                              nrow = 5, 
                              not_as_factor = c("mrn", 
                                                "patient"
                                                ),
                              autoWidth = TRUE
) {
  # mutate(across(where(is.character) using forcats::as_factor, avoiding items in "not_as_factor"
  table <- table |> 
    mutate(across(where(is.character) 
                  & !matches(not_as_factor), 
                  forcats::as_factor))
  
  # configure table settings
  dtable <- DT::datatable(table, 
                          extensions = 'Buttons', 
                          filter = "top",
                          options = list(
                            pageLength = nrow, #nrow dictates pageLength,
                            #lengthMenu = c(5, 10, 15, 20),
                            # initComplete = JS(
                            #   "function(settings, json) {",
                            #   "$(this.api().table().header()).css({'background-color': '#000', 'color': '#fff'});", #we use the initComplete callback function in options to change the background color of the table header to black after the initialization:
                            #   "}"
                            dom = 'Bfrtip', 
                            buttons = c('copy',
                                        'csv',
                                        #'pdf',
                                        'excel'
                                        ),
                            autoWidth = autoWidth,
                            searchHighlight = TRUE)
  )
  
  return(dtable)
}


```

# Overview

This report summarises themes identified from digital requests using locally-hosted LLM-generated metadata.

The report is intended to support prioritisation and portfolio review by:

* identifying common request themes
* highlighting potentially related or duplicate requests
* summarising requests by business area
* identifying records with quality assurance findings

## Summary statistics

```{r}
summary_tbl <- tibble(
  Metric = c(
    "Requests analysed",
    "Business areas",
    "Themes",
    "Requests with QA findings"
  ),
  Value = c(
    nrow(flat),
    n_distinct(flat$business_area, na.rm = TRUE),
    n_distinct(themes$themes, na.rm = TRUE),
    sum(flat$n_findings_v2 > 0 | flat$n_findings_v3 > 0)
  )
)

summary_tbl |>
  flextable() |>
  theme_ft()
```

# Most common themes

## Word cloud

```{r}
#| out-width: 100%
#| fig-width: 16
#| fig-height: 12

theme_counts <- themes |>
  count(themes, sort = TRUE) |>
  #slice_head(n = 75) |> 
  filter(n >= 2)

theme_counts |>
  slice_head(n = 75) |>
  ggplot(
    aes(
      label = themes,
      size = n,
      colour = n
    )
  ) +
  ggwordcloud::geom_text_wordcloud_area(
    eccentricity = 0.5
  ) +
  scale_size_area(
    max_size = 60
  ) +
  scale_colour_viridis_c() +
  theme_void() +
  theme(
    legend.position = "none"
  )
```

## Top 20 themes

```{r}
theme_counts |>
  slice_head(n = 20) |>
  ggplot(
    aes(
      x = reorder(themes, n),
      y = n
    )
  ) +
  geom_col() +
  coord_flip()
```

# Potentially related requests

Requests that share multiple themes may represent:

* similar problems
* opportunities for collaboration
* duplicate requests
* opportunities to consolidate work

```{r}
pairs <- inner_join(
  themes,
  themes,
  by = "themes",
  relationship = "many-to-many"
) |>
  filter(
    digital_request_number.x < digital_request_number.y
  ) |>
  group_by(
    digital_request_number.x,
    digital_request_number.y
  ) |>
  summarise(
    shared_themes = paste(
      sort(unique(themes)),
      collapse = "; "
    ),
    n_shared_themes = n(),
    .groups = "drop"
  ) |>
  filter(n_shared_themes >= 2) |>
  arrange(
    desc(n_shared_themes),
    digital_request_number.x,
    digital_request_number.y
  )

pairs |>
    interactive_table(nrow = 5, 
                                      not_as_factor = c("mrn", 
                                                        "patient"
                                                        ),
                                ) 
```

# Request summaries

```{r}
flat |>
  transmute(
    digital_request_number,
    business_area,
    themes = map_chr(
      themes,
      paste,
      collapse = "; "
    ),
    summary = str_trunc(
      summary,
      width = 1000
    )
  ) |>
  arrange(
    business_area,
    digital_request_number
  ) |>
  interactive_table(nrow = 5, 
                                      not_as_factor = c("themes", 
                                                        "summary"
                                                        ),
                                ) 
```

# Quality assurance findings

The records below have one or more quality assurance findings generated during metadata extraction.

```{r}
qa <- flat |>
  transmute(
    digital_request_number,
    n_findings = n_findings_v2 + n_findings_v3,
    findings = pmap_chr(
      list(findings_v2, findings_v3),
      \(v2, v3) {
        c(v2, v3) |>
          unique() |>
          paste(collapse = "\n\n• ")
      }
    )
  ) |>
  mutate(
    findings = paste0("• ", findings)
  ) |>
  filter(n_findings > 0) |>
  arrange(desc(n_findings))

qa |>
   interactive_table(nrow = 5, 
                                      not_as_factor = c("findings", 
                                                        "patient"
                                                        ),
                                ) 
```

# Appendix

## Business area frequencies

```{r}
flat |>
  count(
    business_area,
    sort = TRUE
  ) |>
  rename(
    Requests = n
  ) |>
  flextable() |>
  theme_ft()
```


